Death by Dropdown? Engineer Insightful Shiny Apps with Behavioral Science

### 2025-09-10 conference-materials/2025_posit-conf/draft.qmd
Refine presentation for readability and accessibility
- Added consistent margin containers (`<div style="margin: 5% 10%;">`) for better layout alignment across slides
- Simplified phrasing in bullet points for easier verbal delivery
- Updated image assets and replaced placeholders with final visuals
- Adjusted type scale and spacing in CSS to ensure minimum font sizes and line heights meet accessibility standards
- Improved contrast and accent consistency for better WCAG compliance

## Dashboard Support Group {.center background-color="#2c3e50"}

:::: {.columns}
::: {.column width="50%" .text-xl}
### Quick show of hands:
- Who's built a Shiny app with 10+ filters?
- Who's had a user ask for "just one more dropdown"?
- Whose dashboard has more `selectInput()` than insights?
:::

@@

@@

:::: {.columns}
::: {.column width="50%"}
**PhD in Applied Psychology**
*How people actually process information*

**Led Shiny teams at scale**
*Built dashboards people use*

**Created {bidux} package**
*Your behavioral scientist in the console*
:::

@@

::: {.text-box-overlay style="position: absolute; left: 50px; top: 150px; width: 600px;"}
### Real client. Real dashboard. Real failure.

Their "comprehensive" solution:
- 18 filters across 6 tabs
- Every possible data view
- Zero user adoption

@@

@@

- **Cognitive Load Theory**: Too many choices → paralysis
- **Progressive Disclosure**: Reveal complexity gradually
- **Framing Effects**: Same data, different story
- **Choice Architecture**: Guide without restricting
:::

@@

:::: {.columns}
::: {.column width="50%"}
### Think of it like a road trip

@@

::: {.column width="50%" .fragment}
### The BID Framework

📍 **Interpret** user needs<br>
⚠️ **Notice** friction points<br>
🌦️ **Anticipate** cognitive biases<br>
🗺️ **Structure** information flow<br>
✅ **Validate** understanding
:::
::::

![](img/bid-framework.png){.absolute bottom="50" right="50" width="650"}

@@

- ✅ Works with ANY Shiny dashboard (existing or new)
- ✅ Analyzes telemetry OR works without it
- ✅ Auto-suggests improvements based on behavioral science
- ✅ Lets you override with custom parameters
:::

::: {.fragment .text-xl}
**Think of it as your dashboard's therapist**
:::

@@

::: {.column width="50%"}
![Marketing: "75% to goal!" 🎉](img/framing-positive-placeholder.png){style="margin-bottom: 20px;"}

![Finance: "25% gap!" 😱](img/framing-negative-placeholder.png)
:::
::::

@@

@@

:::: {.columns style="text-align: center;"}

::: {.column width="33%" .text-display}
**Dashboards don't need more features.**<br>
**They need fewer obstacles.**

@@

::: {.column width="33%" .fragment}
*Special thanks to the posit::conf organizers,<br>
the Shiny community,<br>
and all of you for caring about your users*
:::

::: {.column width="33%"}
![](img/hex-bidux.png){class="hex-logo" width="250"}
:::

::::

>>>>

## Dashboard Support Group {background-color="#2c3e50"}

<div style="margin: 5% 10%;">

:::: {.columns}
::: {.column width="50%" .text-xl}
### Quick show of hands:
- Built Shiny app 10+ filters?
- User asked "one more dropdown"?
- More `selectInput()` than insights?
:::

@@

</div>

@@

<div style="margin: 5% 10%;">

:::: {.columns}
::: {.column width="50%"}
**PhD in Applied Psychology**<br>
*How people actually process information*

**Led Shiny teams at scale**<br>
*Built dashboards people use*

**Created {bidux} package**<br>
*Your behavioral scientist in the console*
:::

@@

::: {.absolute left="-50" bottom="200"}
### Real client. Real dashboard. Real failure.

- 18 filters across 6 tabs
- Every possible data view
- Zero user adoption

@@

<div style="margin: 5% 10%;">

@@

- **Cognitive Load**: Too many choices
- **Progressive Disclosure**: Reveal complexity gradually
- **Framing Effects**: Same data, different story
- **Choice Architecture**: Guide without restricting
:::

</div>

@@

<div style="margin: 5% 10%;">

:::: {.columns}
::: {.column width="40%"}
### Think of it like a road trip

@@

::: {.column width="40%" .fragment}
### The BID Framework

📍 **Interpret** user needs<br>
⚠️ **Notice** friction points<br>
🌦️ **Anticipate** cognitive biases<br>
🗺️ **Structure** information flow<br>
✅ **Validate** understanding
:::

::: {.column width="20%"}
![](img/bid-framework.png){width="100%"}
:::
::::

</div>

@@

- ✅ Works with ANY Shiny dashboard
- ✅ Analyzes telemetry OR works without
- ✅ Auto-suggests behavioral science improvements
- ✅ Custom parameter overrides
:::

::: {.fragment .text-xl .center}
**Think of it as your dashboard's therapist**
:::

</div>

@@

::: {.column width="50%"}
![Marketing: 🎉](img/ui-gain-frame.png){style="margin-bottom: 20px;"}

![Finance: 😱](img/ui-loss-frame.png)
:::
::::

@@

::: {.fragment}
![](img/ui-summary.png){.absolute bottom="50" right="100" width="650"}
:::

</div>

@@

<div style="margin: 5% 10%;">

:::: {.columns .center}

::: {.column width="40%" .text-display}
**Dashboards don't need more features.**<br>
**They need fewer obstacles.**

@@

::: {.column width="30%" .fragment}
*Special thanks to the posit::conf organizers,<br>
the Shiny community,<br>
and all of you for caring about your users*
:::

::: {.column width="30%"}
![](img/hex-bidux.png){class="hex-logo" width="200"}
:::
::::

</div>
